---
title: "Correlation Analysis - Analytes for each Station"
author: "Dave Bosworth"
date: "December 12, 2019"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")
```

# Import and Prepare Data

Load R packages necessary for this analysis:
```{r load packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggforce)
library(corrr)
library(knitr)
library(kableExtra)
library(openwaterhg)
```

Bring in and bind all concentration data:
```{r import and clean conc data, message = FALSE, warning = FALSE}
# Import concentration data
source("YB_Mass_Balance/Concentrations/Import_Conc_Data.R")

# Clean MeHg and THg on solids data
hg_solids_clean <- comb_param_calc %>% 
  filter(Parameter %in% c("THg Concentration on Solids", "MeHg Concentration on Solids")) %>% 
  rename(
    Analyte = Parameter,
    Conc = Value
  )

# Bind all concentration data
all_conc <- bind_rows(all_conc, hg_solids_clean)

# Create a df of all analytes to include and their shortened names
analytes <- tibble(
  analyte_long = c(
    "Chloride- filtered",
    "DOC",
    "Iron- filtered",
    "Manganese- filtered",
    "MeHg- filtered",
    "MeHg- particulate",
    "MeHg- total",
    "POC",
    "Sulfate- filtered",
    "THg- filtered",
    "THg- particulate",
    "THg- total",
    "TOC",
    "TSS",
    "UVA 254",
    "VSS"
  ),
  analyte_short = c(
    "Cl",
    "DOC",
    "fFe",
    "fMn",
    "fMeHg",
    "pMeHg",
    "tMeHg",
    "POC",
    "SO4",
    "fTHg",
    "pTHg",
    "tTHg",
    "TOC",
    "TSS",
    "UVA254",
    "VSS"
  )
)

# Filter and Clean all Conc Data
all_conc_clean <- all_conc %>% 
  # Keep only necessary data
  filter(
    !str_detect(StationName, "Low Flow|^Sac"),
    Analyte %in% analytes$analyte_long
  ) %>% 
  # Shorten Analyte names
  left_join(analytes, by = c("Analyte" = "analyte_long")) %>% 
  # Keep only necessary variables
  select(
    StationName,
    SampleDate,
    analyte_short,
    Conc
  )

# Clean up
rm(analytes)
```

Setup factor order:
```{r set factor orders}
# Stations
sta_order <- c(
  "Fremont Weir- East Side",
  "Fremont Weir- Middle",
  "Fremont Weir- West Side",
  "CCSB Overflow Weir- North",
  "CCSB Overflow Weir- South",
  "Knights Landing Ridge Cut",
  "Putah Creek at Mace Blvd",           
  "Toe Drain at County Road 22",
  "Toe Drain at Interstate 80",
  "Toe Drain at Lisbon Weir",
  "Toe Drain at 1/2 Lisbon",
  "Liberty Cut below Stairsteps",
  "Shag Slough below Stairsteps",
  "Prospect Slough", 
  "Cache Slough near Ryer Island",
  "Miner Slough near Sac River"
)

# Analytes
ana_order <- c(
  "tTHg",
  "fTHg",
  "pTHg",
  "tMeHg",
  "fMeHg",
  "pMeHg",
  "TSS",
  "VSS",
  "TOC",
  "DOC",
  "POC",
  "UVA254",
  "fFe",
  "fMn",
  "Cl",
  "SO4"
)

all_conc_clean <- all_conc_clean %>% 
  mutate(
    StationName = factor(StationName, levels = sta_order),
    analyte_short = factor(analyte_short, levels = ana_order)
  )

# Clean up
rm(ana_order, sta_order)
```

# Create Plot and Correlation Analysis Functions

Function for scatterplot matrices:
```{r function scatterplot matrix}
plot_scatt_matrix <- function(df, PlotSubT) {
  p <- 
    ggplot(
      data = df, 
      aes(
        x = .panel_x, 
        y = .panel_y
      )
    ) +
    geom_point(size = 0.5, na.rm = TRUE) +
    facet_matrix(
      vars(everything()), 
      layer.diag = FALSE
    ) +
    labs(
      title = "Scatterplot Matrix",
      subtitle = PlotSubT
    ) +
    theme(axis.text.x = element_text(angle = 90))
  
  return(p)
}
```

Function to create correlation matrix:
```{r function correlation matrix}
corr_matrix <- function(df) {
  df_corr <- correlate(df, quiet = TRUE)
  
  return(df_corr)
}
```

Function for correlation plots:
```{r function correlation plots}
plot_corr <- function(df_corr, PlotSubT) {
  p <- rplot(df_corr, print_cor = TRUE) + 
    labs(
      title = "Correlation Plot- Pearson's correlation coefficients",
      subtitle = PlotSubT
    ) +
    theme(axis.text.x = element_text(angle = 90))
  
  return(p)
}
```

Function to reformat correlation matrices to publication-quality:
```{r function reformat correlation matrix}
reformat_corr_matrix <- function(df_corr) {
  df_corr_reformat <- stretch(df_corr, remove.dups = TRUE, na.rm = TRUE) %>% 
    pivot_wider(names_from = y, values_from = r)
  
  return(df_corr_reformat)
}
```

Function to calculate p-values for correlation analyses:
```{r function calc pvalues}
calc_corr_pval <- function(df, df_corr) {
  
  # Calculate sample sizes of each pair
  df_N <- crossprod(as.matrix(!is.na(df))) %>% 
    as_tibble(rownames = "x") %>% 
    pivot_longer(
      cols = -x,
      names_to = "y",
      values_to = "Num"
    )
    
  # Add sample sizes to df_corr df and calculate p-values
  pval_matr <- df_corr %>% 
    stretch(remove.dups = TRUE, na.rm = TRUE) %>% 
    left_join(df_N, by = c("x", "y")) %>% 
    mutate(
      t_stat = r * sqrt(Num - 2)/sqrt(1 - r^2),
      p_value = if_else(
        t_stat >= 0,
        2*pt(t_stat, df = Num-2, lower.tail = FALSE),
        2*pt(t_stat, df = Num-2)
      )
    ) %>% 
    select(x, y, p_value) %>%
    pivot_wider(names_from = y, values_from = p_value) 

  return(pval_matr)
}
```

Function to apply conditional formatting to p-value matrices:
```{r function cond format pvalues}
cond_format_pval <- function(pval_matr) {
  pval_matr_cond_format <- pval_matr %>% 
    pivot_longer(
      cols = -x, 
      names_to = "y", 
      values_to = "pval"
    ) %>% 
    filter(!is.na(pval)) %>% 
    mutate(
      pval = signif(pval, 2),
      pval = cell_spec(
        pval, 
        "html", 
        color = if_else(pval < 0.05, "red", "black"),
        bold = if_else(pval < 0.05, TRUE, FALSE)
      )
    ) %>% 
    pivot_wider(names_from = y, values_from = pval)
  
  return(pval_matr_cond_format)
}
```

# Run Correlation Analysis for each Station and Create Plots

Create a nested dataframe grouped by Station Name:
```{r create nested df by StationName}
all_conc_nestdf_sta <- all_conc_clean %>%
  arrange(analyte_short) %>% 
  pivot_wider(names_from = analyte_short, values_from = Conc) %>% 
  select(-SampleDate) %>% 
  group_nest(StationName)
```

Run the plot and correlation functions on nested dataframe:
```{r apply corr and plot functions}
all_conc_corr_analysis <- all_conc_nestdf_sta %>% 
  mutate(
    plots_scatt = map2(data, StationName, .f = plot_scatt_matrix),
    corr_matr = map(data, .f = corr_matrix),
    plots_corr = map2(corr_matr, StationName, .f = plot_corr),
    corr_matr_reformat = map(corr_matr, .f = reformat_corr_matrix),
    pval_matr = map2(data, corr_matr, .f = calc_corr_pval),
    pval_matr_cond_format = map(pval_matr, .f = cond_format_pval)
  )
```

# Results {.tabset .tabset-pills}

```{r print plots and corr matrices, echo = FALSE, message = FALSE, results = "asis", fig.height = 9, fig.width = 13}
for (i in 1:nrow(all_conc_corr_analysis)) {
  # Create subheadings for each Station
  cat("## ", as.character(all_conc_corr_analysis$StationName[i]), " {.tabset .tabset-pills}") 
  cat("\n\n### Plots\n\n")
  # Print Scatterplot Matrices
  print(all_conc_corr_analysis$plots_scatt[[i]])
  cat("\n\n")
  # Print Correlation Plots
  print(all_conc_corr_analysis$plots_corr[[i]])
  cat("\n\n")
  # Print Correlation Matrices
  k1 <- 
    kable(
      all_conc_corr_analysis$corr_matr_reformat[[i]], 
      digits = 2, 
      caption = paste(
        "Pearson's correlation coefficients:", 
        as.character(all_conc_corr_analysis$StationName[i])
      )
    ) %>% 
    kable_styling("striped")
  cat("### Correlation Matrix\n\n", k1, "\n\n")
  # Print p-value Matrices
  k2 <- 
    kable(
      all_conc_corr_analysis$pval_matr_cond_format[[i]],
      format = "html",
      escape = FALSE,
      caption = paste(
        "p-values for correlation coefficients:", 
        as.character(all_conc_corr_analysis$StationName[i])
      )
    ) %>%
    kable_styling("striped")
  cat("### P-value Matrix\n\n", k2, "\n\n")
}
```

# Observations

Significant Correlations:

* Total/particulate THg and TSS were always positive and significant across all stations.
* With a few exceptions, this was mostly the case with filtered THg and TSS as well.
* VSS also had strong and and significant positive associations with all fractions of THg.
* Most stations had significant and positive correlations between filtered Fe and THg; however, some of these relationships had outliers with high leverage. Total and particulate THg were more commonly correlated with Fe than the filtered fraction.

</br>
Insignificant Correlations:

* For the most part, all fractions of MeHg were not correlated with TSS, VSS, all fractions of THg, and filtered Fe.
* Also for the most part, filtered Mn was not correlated with all fractions of THg and MeHg.

</br>
Mixed Results:

* At some of the stations the correlations between TOC/DOC/POC and THg/MeHg were significant and positive, but this was inconsistent.

